Secure Composition of Robust and Optimising Compilers

### 2023-06-16 popl/main.tex
Overview. Combined Monitor proof.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
      \node (S) {$\src{L_{\tmssafe}}$};
      \node[right=1.0 of S] (T) {$\trg{L}$};
      \node[right=1.0 of T] (M) {$\irl{L_{\mssafe}}$};
      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
      \node[right=1.0 of D0] (D1) {$\irl{L_{\mssafe}}$};
      \node[right=1.0 of C0] (C1) {$\irl{L_{\mssafe}}$};
      \node[right=4.5 of M] (O) {$\obj{L_{\scctsafe}}$};

      \draw[->] (S) to[sloped,above] node {\text{\scriptsize\gls{tms}}} (T);
      \draw[->] (T) to[sloped,above] node {\text{\scriptsize\gls{sms}}} (M);
      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (D0);
      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (C0);
      \draw[->] (D0) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (D1);
      \draw[->] (C0) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (C1);
      \draw[->] (D1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
      \draw[->] (C1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
    \end{tikzpicture}
  \end{center}
  \caption{Visualisation of the optimizing compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}
\end{figure}

@@

\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
This section defines several secure compilers, each of which robustly preserves a different property of interest.
The power of the frameworks from \Cref{sec:compprop,sec:compcomp} is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
Concretely, the section starts starts by defining languages $\src{L_{TMS}}$ and $\trg{L}$ and describes a compiler $\cc{\src{L_{TMS}}}{\trg{L}}$ for the robust preservation of \gls{tms} between the two languages (\Cref{subsec:cs:tms}).
Then, it defines $\irl{L_{MS}}$ and a compiler $\cc{\trg{L}}{\irl{L_{MS}}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
By the framework presented in \Cref{sec:compcomp,sec:compprop}, a compiler $\rtp{\cc{\src{L_{TMS}}}{\trg{L}}\circ\cc{\trg{L}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
Two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ are defined that perform \gls{dce} and \gls{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
Finally, the paper defines $\obj{L_{CCT}}$ and a compiler $\cc{\irl{L_{MS}}}{\obj{L_{CCT}}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}).
Composing it with $\cc{\src{L_{TMS}}}{\irl{L_{MS}}}$ and the optimizing compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.

\MKin{add image and do introduction using that}

>>>>

@@

\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
      \node (S) {$\src{L_{\tmssafe}}$};
      \node[right=1.0 of S] (T) {$\trg{L}$};
      \node[right=1.0 of T] (M) {$\irl{L_{\mssafe}}$};
      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
      \node[right=1.0 of D0] (D1) {$\irl{L_{\mssafe}}$};
      \node[right=1.0 of C0] (C1) {$\irl{L_{\mssafe}}$};
      \node[right=4.5 of M] (O) {$\obj{L_{\scctsafe}}$};

      \draw[->] (S) to[sloped,above] node {\text{\scriptsize\gls{tms}}} (T);
      \draw[->] (T) to[sloped,above] node {\text{\scriptsize\gls{sms}}} (M);
      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (D0);
      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (C0);
      \draw[->] (D0) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (D1);
      \draw[->] (C0) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (C1);
      \draw[->] (D1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
      \draw[->] (C1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
    \end{tikzpicture}
  \end{center}
  \caption{Visualisation of the optimizing compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}\label{fig:pipeline}
\end{figure}

This section defines several secure compilers, each of which robustly preserves a different property of interest (see \Cref{fig:pipeline}).
The power of the framework (\Cref{sec:compprop,sec:compcomp}) is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
Concretely, the section starts with a base-line definition for the used languages and extends that to define languages $\src{L_{TMS}}$ and $\trg{L}$.
For programmers using $\src{L_{TMS}}$, their mental model should be that their code is temporal memory safe and cryptographic constant time.
The former is enforced by the static semantics of $\src{L_{TMS}}$, where each well-typed program is temporal memory safe.
The latter is completely untyped.
Having languages defined, this section continues with a compiler $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ and proves that $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ robustly preserves \gls{tms} (\Cref{subsec:cs:tms}).
For readability purposes, it defines $\irl{L_{MS}}$ to be exactly the same as $\trg{L}$ and continues to present a compiler $\cc{\trg{L}}{\irl{L_{MS}}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
Now leaveraging the power of the framework that this paper introduces (\Cref{sec:compcomp,sec:compprop}), a compiler $\rtp{\cc{\src{L_{TMS}}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
The next steps in the compilation pipeline are two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ that perform \gls{dce} and \gls{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
Finally, this section defines the language $\obj{L_{\scctsafe}}$ that is, in contrast to the other languages, not cryptographic constant time by definition.
Concluding, a compiler $\cc{\irl{L_{MS}}}{\obj{L_{CCT}}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}) is shown and composed with $\cc{\src{L_{TMS}}}{\irl{L_{MS}}}$ and the optimizing compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$.
In turn, this yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
